---
title: "Organize Data Notebook"
---

```{r}
#| label: setup
#| include: false
library(here)
source(here("utils","check_packages.R"))
source(here("utils","functions.R"))
```

I set the parameter for the limit on marriage duration I will use to identify new marriages right at the top.

```{r}
#| label: mar-length-param
years_mar <- 5
```

```{r}

#| label: coding-functions

# since I have to code the same variable for both spouses, it makes sense to put
# this coding into a function

code_race <- function(raced, hispand) {
  # We want to take the raced and hispand variables and code them into a combined
  # race variable. I am going to use the fullest possible coding here. 
  race <- case_when(
    # Latino trumps everything, because combined question, sigh
    hispand>0 & hispand<900 ~ "Latino",
    # The original ethnoracial triangle - white, black, indigenous
    raced==100 ~ "White",
    raced==200 ~ "Black",
    (raced>=300 & raced<400) ~ "Indig",
    # Asian
    # need to do South Asians, then Pac Islander, then E&SE Asian
    # to simplify coding nightmare
    # leave out South Asian separation, because we don't have details
    # for those that check more than one
    #(raced %in% c(610, 664, 669, 670)) ~ "South Asian",
    (raced==630 | (raced>=680 & raced<=699)) ~ "Indig",
    ((raced==400 & raced<=679) | raced %in% c(869, 887)) ~ "Asian",
    # Multiracial groups
    raced==801 ~ "White/Black",
    raced==802 ~ "White/Indig",
    (raced>=810 & raced<=819) ~ "White/Asian",
    (raced>=820 & raced<=827) ~ "White/Indig",
    raced==830 ~ "Black/Indig",
    (raced>=831 & raced<=838) ~ "Black/Asian",
    (raced>=840 & raced<=842) ~ "Black/Indig",
    (raced>=850 & raced<=854) ~ "Indig/Asian",
    raced==855 ~ "Indig",
    (raced>=860 & raced<=868) ~ "Indig/Asian",
    #raced==901 ~ "White/Black/AIAN",
    TRUE ~ NA_character_
  )
  
  race <- factor(race, 
                 levels=c("White","Black","Indig","Asian","Latino",
                          "White/Black","White/Indig","White/Asian",
                          "Black/Indig","Black/Asian","Indig/Asian"))
  
  return(race)
}


code_educ <- function(educd) {
  # We want to re-code the educd variable into the following simple
  # categories:
  # - Less than high school diploma
  # - High school diploma
  # - Some college, but less than a four year degree
  # - Four year college degree or more
  educ <- ifelse(educd<60, "LHS",
                 ifelse(educd<=65, "HS",
                        ifelse(educd<=90, "SC",
                               ifelse(educd<=999, "C", 
                                      NA))))
  educ <- factor(educ,
                 levels=c("LHS","HS","SC","C"),
                 ordered=TRUE)
  return(educ)
}

code_bpl <- function(bpld) {
  #recode any one born in the US (99 or less) as a single number. Otherwise
  #we will be fitting state level endogamy. also code in missing values
  
  # The general codes are way to general. The detailed codes have some
  # overspecifity in the codes, but not in the actual data used here.
  return(ifelse(bpld>=95000, NA, 
                ifelse(bpld<10000,1,bpld)))
}

code_language <- function(language) {
  #I need to use the detailed language codes as the general language codes 
  #are too general. However the detailed language codes are too detailed in 
  #some places, particularly in translating between the two time periods. Thus
  #I make some corrections to the detailed codes for consistency between the
  #two time periods.
  
  # We don't need to code language as a factor variable with labels because
  # ultimately all we need to know is whether the two partners speak the same
  # language. We should however code in any missing values (0), although 
  # there do not appear to be any.
  
  lang_recode <- language
  
  #the following cases will be collapsed to the their top two digit codes
  #(starting at the 100 levels)
  #1:27 - European language groups (e.g. English, French, German)
  #35: Uralic
  #37: Other Altaic
  #43: Chinese
  #47: Thai/Siamese/Lao (not separable in 1980)
  #52: Indonesian
  #53: Other Malay
  #58:  Near East Arabic Dialect
  collapsed_cases <- c(1:24,35,37,43,47,52,53,58)
  collapsed_language <- floor(language/100)
  lang_recode <- ifelse(collapsed_language %in% collapsed_cases,
                        collapsed_language*100, lang_recode)
  
  #A couple of cases need to be put back into there other categories
  #420: Afrikaans
  #1140: Haitian Creole
  #1150: Cajun
  #2310: Croatian
  #2320: Serbian
  uncollapsed_cases <- c(420,1140,1150,2310,2320)
  lang_recode <- ifelse(language %in% uncollapsed_cases,
                        language, lang_recode)
  
  #put malay and other malay together
  lang_recode <- ifelse(language==5270, 5300, lang_recode)
  
  #collapse Hindi and Urdu into 3101 (Hindustani)
  lang_recode <- ifelse(language>3101 & language<=3104, 3101, lang_recode)
  
  #For 1980 consistency put all American Indian languages in one group
  lang_recode <- ifelse(lang_recode>7000 & lang_recode<=9300, 7000, 
                        lang_recode)
  
  #A few cases are "other" or "nec". These will be recoded as -1 and
  #not treated as endogamy with each other
  nec_codes <- c(3140,3150,3190,5290,5590,6200,6390,6400,9400,9410,9420,9500,
                 9600,9601,9602,9999)
  lang_recode <- ifelse(lang_recode %in% nec_codes, -1, lang_recode)
  
  #code any missing values 
  lang_recode <- ifelse(lang_recode==0, NA, lang_recode)
  
  return(lang_recode)
  
}
```

Here are the starting and ending values for the ACS 2021 data from the IPUMS codebook. All of the other extracts should be identical in structure.

```
  Variable                         Columns        Len   2021    
  YEAR                         H   1-4            4     X       
  SAMPLE                       H   5-10           6     X       
  SERIAL                       H  11-18           8     X       
  CBSERIAL                     H  19-31          13     X       
  HHWT                         H  32-41          10     X       
  CLUSTER                      H  42-54          13     X       
  STATEFIP                     H  55-56           2     X       
  MET2013                      H  57-61           5     X       
  STRATA                       H  62-73          12     X       
  GQ                           H  74              1     X       
  PERNUM                       P  75-78           4     X       
  PERWT                        P  79-88          10     X       
  SEX                          P  89              1     X       
  AGE                          P  90-92           3     X       
  MARST                        P  93              1     X       
  MARRNO                       P  94              1     X       
  YRMARR                       P  95-98           4     X       
  RACE                         P  99              1     X       
  RACED                        P 100-102          3     X       
  HISPAN                       P 103              1     X       
  HISPAND                      P 104-106          3     X       
  BPL                          P 107-109          3     X       
  BPLD                         P 110-114          5     X       
  YRIMMIG                      P 115-118          4     X       
  LANGUAGE                     P 119-120          2     X       
  LANGUAGED                    P 121-124          4     X       
  EDUC                         P 125-126          2     X       
  EDUCD                        P 127-129          3     X       
  PERNUM_SP                    P 130-133          4     X       
  SEX_SP                       P 134              1     X       
  AGE_SP                       P 135-137          3     X       
  RACED_SP                     P 138-140          3     X       
  HISPAND_SP                   P 141-143          3     X       
  BPLD_SP                      P 144-148          5     X       
  YRIMMIG_SP                   P 149-152          4     X       
  LANGUAGED_SP                 P 153-156          4     X       
  EDUCD_SP                     P 157-159          3     X          
```

```{r}
#| label: create-cols-info
acs_start <- c(1,11,32,55,57,75,89,90,93,94,95,100,104,110,115,121,127,130,134,135,138,141,144,149,153,157)
acs_end   <- c(4,18,41,56,61,78,89,92,93,94,98,102,106,114,118,124,129,133,134,137,140,143,148,152,156,159)
acs_names <- c("year","serial","hhwt","statefip","metarea","pernum",
               "sex","age","marst","marrno","yrmarr","raced",
               "hispand","bpld","yrimmig","languaged","educd","pernum_sp",
               "sex_sp","age_sp","raced_sp","hispand_sp","bpld_sp",
               "yrimmig_sp","languaged_sp","educd_sp")
tibble(variable=acs_names, start=acs_start, end=acs_end) |>
  gt()
```


The ACS data is split into separate files for size, but has an identical structure, so I can just feed a vector of file names into the `read_fwf` function.

```{r}
#| label: read-data
acs <- read_fwf(fs::dir_ls(here("data","data_raw","acs"), glob = "*.dat.gz"),
                col_positions = fwf_positions(start = acs_start,
                                              end   = acs_end,
                                              col_names = acs_names),
                col_types = cols(.default = "i", hhwt="d"), 
                progress = FALSE)
```

```{r}
#| label: code-variables
acs <- acs |> 
  mutate(
    ## unique ids for each respondent ##
    id = serial*1000000+pernum*10000+year,
    id_sp = ifelse(is.na(pernum_sp),NA,
                    serial*1000000+pernum_sp*10000+year),
    ## sex of respondent ##
    sex = ifelse(sex==1, "Male", ifelse(sex==2, "Female", NA)),
    # marital status of respondent
    marst=factor(marst, levels=1:6, 
                 labels=c("Married, spouse present", "Married, spouse absent",
                          "Separated","Divorced","Widowed","Never married")),
    ## race of respondent and spouse - see details in code_race function ##
    race = code_race(raced, hispand),
    race_sp = code_race(raced_sp, hispand_sp),
    ## education of respondent and spouse - see code_educ function for details ##
    educ = code_educ(educd),
    educ_sp = code_educ(educd_sp),
    ## language of respondent and spouse - see code_lang for details ##
    lang = code_language(languaged),
    lang_sp = code_language(languaged_sp),
    ## marriage market id ##
    # combined state and metro area id to get marriage market. Replace state id
    # with metro id for cases where it is not zero. Multiply by 100 to avoid id
    # collision
    # FIXME: this needs to be implemented correctly
    mar_market = statefip,
    ## marriage duration ##
    dur_mar = ifelse(yrmarr==0, NA, year - yrmarr),
    ## country of birth of respondent and spouse - see code_bpl for details ##
    bpl = code_bpl(bpld),
    bpl_sp = code_bpl(bpld_sp),
    ## years living in USA for the respondent and spouse ##
    yr_usa = year-ifelse(yrimmig==0, NA, yrimmig),
    yr_usa_sp = year-ifelse(yrimmig_sp==0, NA, yrimmig_sp),
    ## age at migration to USA, respondent and spouse ##
    # straightforward, but still get some negative values. Most of these are
    # just -1 values due to calendar year/birth year issues, but a few other 
    # cases at older ages go back as far as -5. 
    age_usa = age-yr_usa,
    age_usa = ifelse(is.na(age_usa) | age_usa<0, 0, age_usa),
    age_usa_sp = age-yr_usa_sp,
    age_usa_sp = ifelse(is.na(age_usa_sp) | age_usa_sp<0, 0, age_usa_sp)
  )
```

```{r}
#| label: check-yourself

if(sum(duplicated(acs$id))>0) {
  stop("Duplicted ids in data")
}
if(sum(duplicated(na.omit(acs$id_sp)))>0) {
  stop("Duplicted spousal ids in data")
}

tapply(acs$dur_mar, acs$marst, mean)
```

Ok, now that I have checked all of the data, lets just restrict to the variables that I need to keep things tidy.

```{r}
acs <- acs |>
  select(id, id_sp, mar_market, sex, marst, dur_mar, race, race_sp, educ, 
         educ_sp, lang, lang_sp, bpl, bpl_sp, yr_usa, yr_usa_sp, age_usa, 
         age_usa_sp)
```

I want to eliminate individuals from the analysis who:
* have been in the USA less than the marriage window
* are currently married longer than the marriage window or who have a marriage duration longer than the current window, if any at all

```{r}
acs <- acs |>
  filter((is.na(yr_usa) | yr_usa>years_mar) & 
           (is.na(yr_usa_sp) | yr_usa_sp>years_mar) &
           (is_single(marst) | dur_mar<=years_mar))
```

```{r}
#| label: create-groups

# create unions, and alternate partners
```

```{r}
#| label: create-fake-unions
years_mar <- 5

markets <- create_unions(acs, years_mar, 20)
```


```{r}
#| label: save-data
save(markets, file=here("data","data_constructed","markets.RData"))
```