---
title: "Blurring the Marriage Market? Contemporary Patterns of Multiracial Marriage"
shorttitle: "Blurring the Marriage Market"
shortauthors: Gullickson and Bratter
abstract: Research on interracial marriage has only begun to incorporate the growing mixed-race population. Using the 2010-2019 pooled American Community Survey, we  explore the likelihood of a range of spousal pairings relative to racial endogamy for multiracial people, while accounting for group size and controls for education, age, and immigration status. A distinguishing feature of marriage for multiracial individuals is the possibility of partial overlap in racial identification - having one component race in common with one’s partner.  We find that exact racial endogamy for many multiracial individuals is relatively quite high, once adjusting for group size, and that partial endogamy through overlap increases the likelihood of a union. Furthermore, partial overlap in racial identification between multiracial and monoracial partners reveals the importance of racial classification regimes determining how multiracial individuals are treated on the marriage market. We find no evidence of a general affinity amongst multiracial individuals who do not share racial ancestry nor do we find that multiracial individuals’ partner choices are less affected by race than monoracial individuals. These patterns have implications for the significance of established racial boundaries and the on-going churning of racial categories even as those categories become more ancestrally complex.
keywords: [Race, Multracial, Interracial, Assortative Mating, Intermarriage]
thanks: Replication materials for this article are available at https://github.com/AaronGullickson/multiracial_intermar.
reference-section-title: References
bibliography: ../bibliography/project.bib
csl: ../bibliography/chicago-parenthetical.csl
format:
  aog-article-pdf:
    keep-tex: true 
    #mainfont: Baskerville 
    #sansfont: Futura-Medium
    fontsize: 12pt   
    geometry: margin=1in
    fig-height: 4     # smaller fig heights make floating easier
    fig-width: 7.5    # set to the (full width - margins) of letter   
    fig-pos: "!t"
    colorlinks: true
    urlcolor: red
  submittable-docx:
    keep-md: true
    fig-format: pdf
    fig-height: 4     # smaller fig heights make floating easier
    fig-width: 7.5    # set to the (full width - margins) of letter 
  #submittable-pdf:
  #   keep-tex: true
  #   titlepage: true
  #   endnotes: false
  #   fig-pos: "!t"
  #   fig-height: 4     # smaller fig heights make floating easier
  #   fig-width: 7.5    # set to the (full width - margins) of letter   
---

```{r}
#| label: setup
#| include: false
library(here)
source(here("utils","check_packages.R"))
source(here("utils","functions.R"))
load(here("data","data_constructed","models.RData"))
load(here("data","data_constructed","markets.RData"))
load(here("data","data_constructed","unions_alts.RData"))

```

```{r}
#| label: create-coef-table
get_coefs <- function(x) {
  temp <- coef(x)[,c(1,2)]
  return(tibble(variable=rownames(temp), coef=temp[,1], se=temp[,2]))
}

coef_short_sym <- get_coefs(model_short_sym) |>
  filter(str_detect(variable, "race_exog")) |>
  mutate(variable=str_replace(variable, "race_exog_short_sym",""))

coef_short_asym <- get_coefs(model_short_asym) |>
  filter(str_detect(variable, "race_exog_short_asym")) |>
  mutate(variable=str_replace(variable, "race_exog_short_asym",""))

coef_full <- get_coefs(model_full_sym_multi) |>
  filter(str_detect(variable, "race_exog_full")) |>
  mutate(variable=str_replace(variable, "race_exog_full_sym",""))

coef_multi <- coef_short_sym |>
  filter(str_detect(variable, "Multi/Multi"))
```

# Introduction

Race is a particularly powerful shaper of mate selection. While marrying across racial lines has increased substantially over the last half century in the United States [@gullickson_black_2006a; @livingston_intermarriage_2017a], the tendency for partners to match on race continues to exceed that of matching on education, religion, or other achieved or ascribed characteristics [@lundquist_interracial_2024; @rosenfeld_racial_2008b]. Understanding the ways racial “sorting” occurs when selecting a partner provides a window into the ongoing salience of racial boundaries. A lower relative likelihood of partnering across certain racial lines signals more rigid social barriers between those groups, while a greater likelihood indicates that the social boundaries between those groups may be flexible, fluid, or even dissolving [@qian_social_2007a; @alba_remaking_2003; @kao_company_2019a]. Moreover, a higher likelihood of interracial partnering for some racial pairings compared to others suggests racialized preferences and on-going social hierarchies [@curington_romantic_2022].

While much scholarly work has established trends and patterns of interracial marriage among groups identifying with one race [@kalmijn_trends_1993a; @qian_changing_2011a; @lundquist_interracial_2024], we have few comprehensive quantitative assessments of interracial marriage among individuals who identify with more than one race. A growing literature has attempted to understand patterns of interracial partnering among multiracial individuals, but this work has either focused on specific racial boundaries or on solely part-white multiracial groups [@alba_great_2020a; @littlejohn_race_2019a; @miyawaki_expanding_2015c]. Other work provides in-depth qualitative exploration of racialized notions shaping multiracial partnering [@song_what_2015a; @song_multiracial_2016a; @buggs_color_2019a;  @mills_colors_2021a].

Multiracial adults are a rapidly expanding segment of the U.S. population. As of 2019, 2.5% of the adult population identified with more than one race (Parker & Barroso 2021). This growth in the number of adults who are both mixed and self-identify as multiracial when partnering complicates notions of boundaries around racial groups and the ancestral complexity within those groups. The need for research in this area is particularly notable given the clear connection between interracial partnering in one generation and the production of individuals with multiple racial ancestries in the next. A focus only on the production of “first generation” biracial individuals from the union of two monoracial partners provides an incomplete view of the complexity of multiracial populations [@bratter_will_2007a; @morning_generational_2018a]. @bratter_will_2007a has shown that the children of self-identified multiracial individuals are often identified by the shared single race of their partners. Thus, the long-term growth of the self-identified multiracial population will depend substantially on the partner choices made by self-identified multiracial individuals themselves. 

As a test of the integrative and boundary blurring force of marriage, we ask what role does race play as a sorting mechanism within marriage for multiracial populations? To address this question, we examine patterns of racial assortative mating in marriage for self-identified multiracial individuals in American Community Survey (ACS) data from 2010-2019. We employ conditional logit models to examine the likelihood of marrying across racial lines, which allows us to adjust for differences in group size as well as a variety of control variables. We employ the concept of *partial overlap in racial identification* to understand the complex nature of matching for multiracial individuals whose partners share one but not both races for marriages involving both multiracial to monoracial couples and multiracial to multiracial couples.

This study uses self-identification on the race question to identify multiracial individuals, which is only one of the multiple ways that multiraciality can be conceptualized and measured [@goldstein_multiplerace_2000a; @roth_multiple_2016c]. Classification on administrative forms is a type of public identification that provides a window into the ways racial identities are experienced in interactions, reflecting how one is or desires to be viewed by outsiders [@brunsma_interracial_2005a; @davenport_role_2016a; @roth_multiple_2016b]. For this study, self-identification serves the dual purposes of allowing us to better understand both the multigenerational processes governing the persistence or disappearance of multiracial identification and identifies the population that is most visibly negotiating being multiracial on the marriage market [@buggs_color_2019a; @mills_colors_2021a].

# The Complexity of Multiracial Partnering

Research on racial assortative mating makes a primary distinction between endogamy – marrying within one’s own group – versus exogamy – marrying outside of one’s own group [@lundquist_interracial_2024]. Racial endogamy indicates a degree of social closure around racial groups and the rigidness of race as a social boundary, while racial exogamy indicates that racial boundaries are easier to cross, more flexible, and less salient [@alba_great_2020a; @kao_company_2019a; @lee_diversity_2010a].

Much of what we know about racial assortative mating neglects marriages involving multiracial adults [@lundquist_interracial_2024]. By virtue of their very nature, multiracial populations complicate the clear distinction between endogamy and exogamy that drives much of this research [@mills_colors_2021a; @song_multiracial_2017a]. Applying the logic of racial endogamy to the case of multiracial partners, we might observe what could be called exact endogamy, where individuals have a higher propensity to partner with others who share the exact same combination of races (e.g., a White-Asian person marries another White-Asian person). However, two issues complicate such an approach to thinking about endogamy for multiracial individuals. First, group size will play an outsized role in the partner choices of multiracial individuals. Second, unlike single-race individuals, multiracial individuals have the possibility of a partial “overlap” in racial identification with potential partners [@littlejohn_race_2019a; @mills_colors_2021a].

## Opportunities and Preferences: The Role of Group Size

Scholars of assortative mating often make a theoretical distinction between opportunities and preferences when analyzing intermarriage patterns [@harris_how_2005a; @kalmijn_intermarriage_1998a; @schwartz_trends_2013a]. Preferences relate to the specific inclinations that individuals will have to partner with a given group or characteristic, while opportunities refer to the structural conditions that govern people’s exposure to potential partners. Both preferences and opportunities play some role in determining the overall pattern of assortative mating. However, rarely are researchers able to fully disentangle these two concepts in empirical work. Furthermore, the concepts themselves may not be fully distinct as social structure may shape people’s ultimate partner preferences [@schwartz_trends_2013a]. Thus, when interpreting patterns of interracial marriage, we must acknowledge that they reflect both the contribution of opportunity and preferences, which is often characterized more generically as “social distance.”

However, researchers generally do adjust for one clear and obvious component of structural opportunity - group size. Group size, particularly on a local level, is a critical determinant of who marries whom for the simple reason that it determines the relative number of potential partners of a given group [@blau_heterogeneity_1982a; @kalmijn_intermarriage_1998a] . All other things being equal, relatively smaller groups will be less endogamous than larger groups because of the relative scarcity of same-group potential partners. For example, levels of Black/White intermarriage are generally lower in the Southern United States where Black populations are concentrated than in the Midwest or the West [@bratter_multiracial_2017a; @gevrek_interracial_2014a]. Similarly, @choi_marriagemarket_2017a show that the outmarriage rates for all racial groups vary inversely with the relative size of the group in the metropolitan area.

Generally, researchers use modeling techniques that adjust for these differences in group size, treating them as “nuisance” terms. Such adjustment does not fully remove all potential opportunity components from measures of assortative mating (e.g., segregation in neighborhoods and social networks), but does provide a more straightforward measure of social distance. Ultimately, this adjustment is critically important for gauging differences in intermarriage patterns when group size varies. However, such adjusted likelihoods can be very different from actual cross-tabulations of who marries whom when differences in group size are substantial.

For multiracial populations, the role of group size is considerable given the relatively small size of combination-specific multiracial populations as well as the broad variation in group concentration across geographic spaces. In simple terms, even if multiracial individuals strongly prefer to marry individuals of the same exact racial combination as themselves, they would find it difficult to do so, given the scarcity of such potential partners in the marriage market. For example, the White-Asian population is among the largest multiracial groups. Yet, they make up less than 1% of the adult US population [@alba_unappreciated_2024a]. White-Asian individuals will therefore find it challenging to find other eligible White-Asian potential partners. Furthermore, even if White-Asian individuals equally prefer single race White and Asian partners, they will be more likely to partner with White spouses, given the fact that White people constitute the racial majority in the United States [@miyawaki_expanding_2015c; @alba_great_2020a].  

By using methodological techniques that account for group size differences, we can recover any strong tendencies toward endogamy among multiracial populations. However, in actual practice, even a strong tendency toward endogamy is likely to be swamped by group size differences. Recognizing the discrepancy here is important, because both features may be valuable to explore. Adjusting for group size provides us a better sense of social distance, while the actual partnering patterns, which contribute to the racial ancestry of the next generation, are heavily influenced by group size.

## Theorizing Shared Racial Identification for Multiracial Marriage

Conventional notions of exact racial endogamy are insufficient to explain partnering of multiracial individuals (Song 2016; 2015).  We, therefore, consider alternate ways to conceptualize endogamy to account for how racial sameness operates for racially mixed adults.  Partnering by this group may involve a partial overlap in racial identification between spouses as the multiracial person may share a single component race with either a monoracial spouse (e.g. White-Asian/White; Black-Asian/Black) or a multiracial spouse (e.g. White-Asian/White-Black).[^1] We consider this a weaker, or partial, form of racial endogamy, but nonetheless demonstrating the principle of endogamy as individuals are similar on a facet of their racial identification.  Such cases complicate the notion of what it means to intermarry or cross a racial boundary [@littlejohn_race_2019a; @mills_colors_2021a; @song_what_2015a]. @littlejohn_race_2019a theorizes that this represents a “shared racial identity” between spouses which is a core part of how racial commonality operates for relationships involving multiracial people. Such sharing can be mapped on a continuum or a “hierarchy of relatedness” [@dacosta_making_2007a] where strongest or most frequent relationships involve unions of two multiracial persons whose combined racial identification exactly matches are followed by those where a partial overlap in racial identification exists between couples and the least frequent would be unions between those individuals with no overlapping racial identification. Partial overlap would signal a form of endogamy that accounts for the potential for complex racial identities among spouses.

[^1]: We use dashes to combine the identified races of self-identified multiracial groups and slashes to indicate marriage between two different multiracial or monoracial groups. A person who identified as White and Asian would be indicated as “White-Asian” while a marriage between a single-race White person and a single-race Asian person would be indicated as a “White/Asian” marriage. Similarly, a “White-Asian/White” marriage would indicate a marriage between a multiracial person with both a White and Asian identification and a person who identified as White alone.

Do multiracial partners seek overlap for some aspects of their racial identification more so than others? To explore this more broadly, we introduce several scenarios, drawn from the concept of “racial classification regimes” [@gullickson_choosing_2011a; @iverson_regimes_2022a; @xu_gender_2021a]. These regimes refer to the ways some recognized racial ancestries have historically over-determined or dominated racial identification and classification among mixed-race people, while other racial ancestries have a more receding or less enduring impact. While similar to the classification concepts of hypodescent and hyperdescent [@davis_who_1991a], the notion that an ancestry dominates or recedes in influence is not necessarily tied to its social status or social ranking in a racial hierarchy and allows for broader generalization in cases where relative social status or ranking may be unclear. We extend this concept from characterizing classification to describing affinities in partnering. As racial identification is also social, it is linked to the ways individuals see other people as well as how they are seen by others. As @xu_gender_2021a [p. 1610] argue, “With this concept, we aim to encapsulate ways of both thinking about and acting on racial status with reference to mixed race people of varying ancestral combinations.” We envision that such “thinking and acting” also includes how potential romantic partners are viewed. Specifically, we hypothesize that partial overlap in racial identification will more likely lead to a union when the partners share a dominant racial identification.

We consider four classification regimes, ordered by dominance of specific ancestry: Black, Indigenous, White, and Asian/Latino.  The Black classification regime refers to the ways partial Black ancestry generally dominates one’s identity amidst other component racial groups. This regime reflects the historical power of the “one-drop” rule [@davis_who_1991a]. Contemporarily,  partial Black ancestry has a high likelihood of translating to either a “Black” identity or at minimum a strong awareness of one’s Blackness [@khanna_if_2010a; @lee_diversity_2010a; @davenport_role_2016a; @roth_end_2005a].  Indeed, part-Black multiracial individuals are more likely than non-Black multiracial individuals to have Black spouses and cohabiting partners [@littlejohn_race_2019a]. Therefore, we expect partial overlap in Black identification (e.g. Black-Asian/Black or Black-Asian/White-Black) to more strongly predict union formation than when a part-Black multiracial individual has partial overlap in non-Black identification. (e.g. Black-Asian/Asian or Black-Asian/White-Asian).

At the other end of the spectrum is the Indigenous regime, which refers to the fading or “recessive” influence of partial American Indian, Alaska Native, Native Hawaiian, or Pacific Islander ancestry on one’s classification [@iverson_regimes_2022a]. This expectation reflects a similarly long-standing convention of identification where individuals of partial Indigenous heritage are more likely to identify with and be accepted as members of other racial groups [@snipp_american_1989a]. For most of US history, this practice has contributed to Indigenous population decline and the stereotype of the "dying Indian" [@wolfe_land_2001a]. Consistent with this convention, far more Americans are likely to report an American Indian ancestry on Census forms than identify as American Indian by race [@liebler_joining_2016a]. In a slight reversal of the historical norm, the American Indian population has grown largely through greater identification since the mid-Twentieth century [@eschbach_changing_1993a]. However, this growth is concentrated among individuals with the least cultural ties to Indigenous life and who are the most likely to change racial identification over time [@liebler_joining_2016a]. When ties to ethnicity and culture are thinner, shared identification is less likely to influence union formation. Therefore, we expect that part-Indigenous individuals will be more likely to form a union with those whom they share non-Indigenous identification (e.g. White-Indigenous/White or White-Indigenous/White-Black) than Indigenous identification (e.g. White-Indigenous/Indigenous or White-Indigenous/Black-Indigenous).

The remaining classification schemes predict something between complete dominance or recession. This expectation is clearest in the case of White ancestry, which varies strongly depending on the other group in consideration. White ancestry behaves as the recessive ancestry in relation to Black ancestry and as the dominant ancestry in relation to Indigenous ancestry, occupying a middle ground between these two poles [@iverson_regimes_2022a].  

For different reasons, Asian and Latino ancestry may also be flexible and occupy a similar intermediate position, owing to less institutionalized historical norms [@gullickson_choosing_2011a].  These groups immigrated in large numbers after the historical development of racial classification regimes for White, Black, and Indigenous ancestry. Historically, both groups have occupied various and often ambiguous places within the U.S. racial hierarchy [@loewen_mississippi_1988a; @bonilla-silva_biracial_2004a; @fox_defining_2013a]. In the contemporary period, Asian and Latino identification is complicated by intragroup heterogeneity along generational, class, and ethnonational lines [@jimenez_mexican_2008a; @jimenez_how_2015a]. Furthermore,  on-going intergenerational incorporation of these groups suggests relatively less residential and social segregation from the White population, reduced social distance, and more permeable boundaries than other minority groups [@alba_remaking_2003]. 

In general, we expect the partner choices of both part-Asian and part-White multiracial individuals to reflect the influence of the regimes at the poles when those regimes are involved (i.e. dominant in combination with Indigenous identification and recessive in combination with Black identification). In the case of White-Asian respondents, we are comparing two ancestries in the middle, about which we have less prior expectation of relative dominance or recession. Prior work shows substantial partnering of White-Asians with both White and Asian individuals, [@curington_tipping_2020a; @miyawaki_expanding_2015c] suggesting greater flexibility. 

As we discuss further below, the structure of the race question on the American Community Survey prevents our ability to directly examine the experiences of part-Latino respondents. However, given the historic ambiguity of Latino identity, we expect more heterogamous union formation for Latino respondents in general. Furthermore, a different history of race mixing in Latin America [@telles_race_2009a] may encourage greater openness to interracial marriage with mixed-race individuals, irrespective of specific ancestry.

## Beyond Racial Overlap: Regimes of Racial Flexibility

Partnering based on a connection to specific racial ancestry or ancestries may not be the sole pattern of multiracial spouse selection. Rather, multiracial adults' spousal selection may challenge or redefine racial boundaries. First, multiracial individuals may partner with other multiracial individuals simply based on the *shared experience of being multiracial*. This refers to the unique social position of multiracial people whose identification defies institutionalized notions that normalize racial categories as encompassing single-ancestry groups [@harris_critical_2016]. Prior research reveals that a shared mixedness may structure day-to-day experience in a way that could lead to the formation of marriages [@dacosta_making_2007a; @jones_who_2011a]. A motivating force behind a “check all that apply” race question was the recognition of a growing community of racially mixed people [@dacosta_making_2007a; @williams_mark_2006a; @nobles_shades_2000a]. This community draws together mixed people around shared interests [@dacosta_multiracial_2020a; @bratter_theres_2022a], though these can also be based around racially specific combinations (e.g. Hapa organizations). Additionally, a growing literature documents a shared type of racialization that specifically stigmatizes being multiracial or racially ambiguous [@johnston_multiracial_2010a]. This shared experience could translate into a propensity of multiracial individuals to partner with one another without regard to exact or partial overlap in racial identification.

Finally, it is possible that race, or ties to specific racial groups, may not be as salient for the selection of spouses among multiracial partners. Borrowing from the typology of racial identity developed by @rockquemore_black_2008a, multiracial partnering may reflect a *transcendent pattern*, where race is not as clearly a guide for how multiracial individuals sort into relationships. According to this argument, we would expect a low level of racial endogamy and high level of racial exogamy across all multiracial groups. This pattern fits a popular narrative of the “browning of America”, where multiracial individuals are harbingers of a post-racial future [@hochschild_creating_2012a; @jones_making_2023a; @strmic-pawl_eleven_2016a]. To the degree that intermarriage represents an integrative force, multiracial people would embody this force, through enhanced tendencies towards interracial relationships and viewing race as irrelevant to partner selection [@song_what_2015a; @curington_romantic_2022; @qian_social_2007a]. Some accounts reveal that multiracial people actively seek out racial differences when partnering [@buggs_color_2019a]. In general, such a pattern supports an eventual blurring of racial boundaries, where racial categories are on the pathway towards being more difficult to discern [@alba_remaking_2003].


# Data and Methods

Data for the analysis are derived from the American Community Survey (ACS) pooled from 2010 to 2019. The ACS is a 1-in-100 annual survey of the US population, conducted by the US Census Bureau. The data were extracted via the IPUMS system [@ruggles_ipums_2023a]. Due to the relatively small number of some multiracial respondents in each annual survey, we pool data for the full decade to create an analytical dataset. We use the most recent pre-pandemic ten year period of the ACS. Additionally, this span of years employs the same question wording and response options on race and Hispanic origin.

## Sample

The basis for the analysis is all opposite-sex marriages formed in the previous five years. We further limit the sample of both marriages and alternate partners to individuals who have been resident in the US for at least the last five years to ensure that the marriage was formed in the US. The sample therefore includes foreign-born individuals who were married or eligible to be married in the US in the last five years. We focus only on opposite-sex couples because the modeling approach discussed below would require using different alternate partner data and models for same-sex couples. The small size of the same-sex population in combination with the small size of the multiracial population would make it difficult to fit models and provide generalizable claims for same-sex couples.

We choose five years as our marriage duration limit to balance potential issues of bias with the need for a large enough sample size, particularly of multiracial partners. In any study of extant marriages, some marriages will have already dissolved, leading to some bias relative to the entire universe of marriages. Such bias is relevant in the study of interracial marriage because interracial marriages have higher dissolution rates than racially endogamous marriages [@bratter_will_2008a]. This bias can be mitigated by limiting marriages to those of recent duration, which has the added benefit of providing information on recent rather than historical marriage patterns. However, a reduction in duration will also reduce the overall sample size. A five year duration limit still provides information on recent marriages while also providing a sufficient sample size for all parts of our analysis. 

We employ conventional practices combining responses to ACS questions on race and Hispanic origin into five racial categories of White, Black, Asian, Indigenous, and Latino. Respondents who self-identified as Hispanic or Latino, regardless of race, are defined as Latino. We combine the American Indian, Alaskan Native, and Pacific Islander responses together as an Indigenous category. A small number of non-Hispanic some other race respondents are dropped from the analysis. We collapse a substantial amount of ethnonational heterogeneity to create the three panethnic categories of Asian, Indigenous, and Latino. Although exploring this heterogeneity might be valuable, the complexity of model parameters we use grows exponentially as the number of single race categories increases. Therefore, for analytical clarity, we limit the single race categories to this small set.

From these single race categories, we identify six biracial categories of White-Black, White-Asian, White-Indigenous, Black-Asian, Black-Indigenous, and Asian-Indigenous.[^2] We limit the analysis to respondents who were either classified with a single race or as one of these six biracial categories. A small number of respondents identified with more than two races, but these were removed as the number in any specific combination is too small to analyze in the models we use here.

[^2]: In deciding how to order the naming of biracial groups, we rely on overall group size, where the larger single-race group is listed first.

Notably, due to the separate question format for Hispanicity, we cannot identify multiracial part-Latino individuals in a way that parallels the other multiracial groups. This is a common shortcoming of research on multiraciality with data that use the Census-style questions that separate race and Hispanic origin. Although Latinos can (and often do) report multiple races, we do not distinguish Latinos by the number of race responses they provide as it is unclear if these responses convey a partial Latino identity. We do, however, include Hispanic adults as potential partners of multiracial individuals.

Because we measure the race response after the formation of the union, it is possible that some race responses differ from what the respondent would have said prior to the union. Race responses can change over time [@saperstein_racial_2012a] and such fluidity is more common for individuals with mixed race ancestry [@liebler_america_2017b]. We expect that spouses are more likely to shift to the same racial identification after marriage, as historical work has found [@saperstein_mulatto_2013a]. This tendency may be exacerbated by a tendency for one partner to fill out the form for all household respondents, although @whitehead_if_2021 found only a small tendency for marital partners to misidentify their spouse’s preferred race. Such a pattern would artificially bias estimates of endogamy upward and estimates of exogamy downward. Thus, estimates of endogamy in our results may have some artificial upward bias.

Although it is possible to identify co-resident unmarried partners in the ACS data, we have not included these cohabiting unions in our analytical sample. Our reasons for this exclusion are twofold. First, we lack information on the duration of current cohabiting unions, which makes it impossible to precisely target recent cohabitations in a way that is analogous to our procedure for marriages. Second, the inclusion of cohabiting unions would necessitate a comparison between marital and cohabiting unions, which is beyond the scope of what is feasible for the current analysis. Prior work has consistently shown that patterns of endogamy/homogamy in cohabitation generally mirror those in marriage but with less strength [@blackwell_homogamy_2004a]. Thus, an analysis of marriages is more feasible with the data available to us and likely offers the best evidence on the strength of existing boundaries.


## Modeling Approach

We use a conditional logit technique to measure the likelihood of interracial marriage. This technique compares the characteristics of each actual marriage to a set of alternate marriages that did not occur. For each actual marriage, we construct a choice set of seven alternate marriages by sampling from a pool of alternate partners for a randomly chosen spouse. We then use a conditional logit model to examine how the characteristics of the two potential partners predict the likelihood of the actual union as opposed to the counterfactual unions within each choice set [@gullickson_counterfactual_2021b]. This method has been used previously to study patterns of assortative mating [@jepsen_empirical_2002a; @qian_marriage_2018a; @gullickson_patterns_2022a] and has properties similar to a more traditional log-linear model. Like a log-linear model, the conditional logit model accounts for differences in group size when estimating the likelihood of marriage by the use of an alternate partner pool that mirrors those group size differences. Unlike a log-linear model, the conditional logit model has the advantage of being able to identify a pool of alternate partners that more closely resembles the choices that individuals have. We use this feature to identify alternate partners as both unmarried and recently married individuals living in the same local geographic area, as described below.

To create alternate unions, we draw from a pool of alternate partners, which includes all unmarried individuals and individuals married in the last five years, based on the year and geographic area of the actual union. The lowest geographic level available in the ACS are Public Use Microdata Areas (PUMAs), which can vary substantially in geographic size. However, the IPUMS data includes additional identifiers for metropolitan area and county when a PUMA is unambiguously contained within each type of geography. For the majority of respondents, we can identify a metropolitan area and/or county. We therefore define marriage markets using a cascading logic. For 74.3% of cases, we are able to identify the metropolitan area of a respondent and we use this as the marriage market. In an additional 2.6% of cases, we are able to identify and use the county of residence as the marriage market. In the remaining 23.1% of cases, we use the PUMA as the marriage market. 

@tbl-race-counts shows the count of actual spouses and alternate partners by race in the final analytical dataset. In total, we analyze 829,485 actual marriages formed in the last five years. Aside from the Indigenous group, the sample size of actual spouses from each monoracial group is quite large, providing a high level of precision to estimates involving those groups. The sample size of multiracial groups is much smaller, particularly for non-part-White multiracial groups. These smaller sample sizes will be partially offset by the large size of monoracial groups when estimating parameters for monoracial-to-multiracial intermarriage, but estimates of multiracial-to-multiracial intermarriage will suffer from more statistical noise as a result of the smaller sample sizes.

```{r}
#| label: tbl-race-counts
#| tbl-cap: Number of spouses and alternate spouses by sex and race from American Community Survey Data, 2010-2019.

race_order <- c("White", "Black", "Asian", "Indigenous", "Latino", 
                "White-Indigenous", "White-Asian", "White-Black",
                "Black-Indigenous", "Black-Asian", "Asian-Indigenous")

unions <- unions |>
  mutate(raceh = factor(raceh, race_order),
         racew = factor(racew, race_order))

male_alt <- male_alt |>
  mutate(raceh = factor(raceh, race_order))

female_alt <- female_alt |>
  mutate(racew = factor(racew, race_order))

tbl_counts <- unions |> 
  count(raceh) |> 
  rename(race = raceh, n_union_h = n)
tbl_counts <- unions |> 
  count(racew) |> 
  rename(race = racew, n_union_w = n) |>
  full_join(tbl_counts)
tbl_counts <- male_alt |> 
  count(raceh) |> 
  rename(race = raceh, n_alt_h =n) |>
  full_join(tbl_counts)
tbl_counts <- female_alt |> 
  count(racew) |>
  rename(race = racew, n_alt_w =n) |>
  full_join(tbl_counts) |>
  select(race, n_union_w, n_union_h, n_alt_w, n_alt_h)

tbl_counts |>
  gt(rowname_col = "race") |>
  fmt_integer(starts_with("n_")) |>
  cols_label(
    race = "Race",
    n_union_w = "Wife",
    n_union_h = "Husband",
    n_alt_w = "Women",
    n_alt_h = "Men"
  ) |>
  cols_align(race, align = "left") |>
  tab_spanner(
    label = "Marriages",
    columns = starts_with("n_union_")
  ) |>
  tab_spanner(
    label = "Alternates",
    columns = starts_with("n_alt_")
  ) |>
  grand_summary_rows(fns = list(list(label = "Total", fn = "sum")),
                     fmt = ~ fmt_integer(.))
```

In all models, we estimate racial exogamy using dummy variables to indicate a particular form of exogamous union, where exact racial endogamy (marrying someone of the same exact race or combination of races) serves as the reference. The estimates from these variables provide the log odds ratio of such a union relative to a racially endogamous union, when all other variables in the model are held constant.[^3] We classify respondents into eleven single-race and multiracial groups in total, meaning that if we were to consider every possible exogamous combination, including the specific sex combination of the partners, the model would have 110 racial exogamy parameters. In practice, such a model would be very difficult to fit. Therefore, we instead estimate models with more constrained versions of these parameters. First, we generally use gender-symmetric parameters that do not distinguish which partner is which gender (e.g. we use a single parameter to predict unions between a White woman and a Black man and unions between a Black woman and a White man). These models effectively estimate the average racial exogamy across both gender combinations. Second, we create choice sets using a subsample of the full data that contains only multiracial individuals. We use this dataset to estimate gender-symmetric models of the likelihood of multiracial-to-multiracial racial exogamy across all fifteen possible combinations of the six multiracial categories.[^4] Based on these results, we develop a reduced set of parameters to model multiracial-to-multiracial racial exogamy that we then apply to models of the full data that also estimate monoracial-to-monoracial and monoracial-to-multiracial racial exogamy parameters.

[^3]: In practice, these models also require that we control for the race of the alternate partner to fully account for differences in group size.

[^4]: In models with no other covariates, these models applied to the reduced dataset will produce results identical to the full dataset that includes monoracial individuals. In practice, these models may differ slightly because of sampling variability in the partners chosen and because the estimated effect of other covariates will differ in the reduced dataset. However, we strongly prefer this approach due to difficulties both in time of computation and model convergence when estimating these parameters for small groups on the full data.

All models include several other control variables. We control for the age difference and its square between potential spouses. We control for education by using educational crossing parameters applied to a set of four ordinal education categories (less than high school, high school, some college, and college). We also control for whether a potential marriage is educationally hypergamous (husband has more education than wife) or educationally hypogamous (wife has more education than the husband). 

Although we restrict the foreign-born population to individuals who were married or eligible to be married in the United States in the last five years, foreign-born individuals may still have a strong tendency toward endogamy with other members of the sending nation. To account for this endogamy, we add two additional control variables. First, we control for language endogamy with a simple dummy variable indicating whether the partners speak the same primary language.[^language] We control for birthplace endogamy in a similar manner, based on the birthplace of each potential partner. However, consistent with @gullickson_patterns_2022a, we adjust the birthplace endogamy parameters for individuals who arrived in the US before age 13 so that such individuals are considered birthplace endogamous with potential partners who were born either in the US or their actual birthplace. 

[^language]: Like race reporting, language endogamy may be somewhat overstated due to a tendency for partners to begin speaking the same primary language after the union has formed.

In addition to regular sampling variability, the model estimates will vary due to the sampling of alternate partners to construct choice sets. We account for this increased variability by generating five sets of data and applying the models separately on each set. We then pool the results and account for this additional variability in the standard errors as described in @gullickson_counterfactual_2021b.[^numsamples]

[^numsamples]: In selecting the number of samples, we attempt to balance improved estimation with computational efficiency. During the preliminary analysis and model building phase, we used only three sets of data to provide faster estimation. Exploratory analysis revealed that the results were highly consistent when using three, five, or seven samples, although computational time increased exponentially with the number of samples.

# Results

@fig-partner-choice-panel-color shows the simple distribution of the race of multiracial spouses’ partners. Each panel shows the patterns for one of the six multiracial groups. The bars distinguish unions involving multiracial partners where there is a) no overlap in their respective racial identification, b) partial overlap in racial identification, and c) an exact match. These distributions make no adjustment for group size or any other covariates of potential spouses that might drive union formation. The importance of group size in determining partners is apparent by how common White partners are across all six multiracial groups, driven by the relatively large size of the White population in the US. For example, White partners represent 66% of all alternate partners, as shown in @tbl-race-counts. However, we can also see the importance of partial overlap in racial identification. White partners are the most common partner choice for all three part-White multiracial groups, representing more than 50 percent of spouses in all three cases. While they are the most common partner for two of the three non-White multiracial groups as well, they never reach this 50 percent benchmark, and are well below it for both part-Black groups. Nonetheless, the overall effect of a partial overlap in racial identification is hard to determine from @fig-partner-choice-panel-color due to the substantial influence of group size.


```{r}
#| label: fig-partner-choice-panel-color
#| fig-cap: Distribution of partner's race for multiracial spouses, separately for each multiracial group.
#| warning: false

couples <- mar_markets[[1]] |> filter(choice)

husbands <- couples |>
  mutate(race_ego=raceh, 
         race_spouse=racew,
         gender="husband") |>
  select(race_ego, race_spouse, gender)
wives <- couples |>
  mutate(race_ego=racew, 
         race_spouse=raceh,
         gender="wife") |>
  select(race_ego, race_spouse, gender)

partner_choice <- bind_rows(husbands, wives) |>
  mutate(race_spouse=factor(race_spouse, 
                            levels=c("White","Black","Indigenous","Asian",
                                     "Latino","White-Asian", "White-Indigenous",
                                     "White-Black", "Black-Asian",
                                     "Black-Indigenous", "Asian-Indigenous")))

partner_choice |>
  filter(str_detect(race_ego, "-")) |>
  group_by(race_ego, race_spouse) |>
  summarize(n = n()) |>
  pivot_wider(id_cols = race_ego, names_from = race_spouse, values_from = n,
              names_prefix = "n_") |>
  ungroup() %>%
  mutate(total = rowSums(select(., starts_with("n_")))) |>
  pivot_longer(cols=starts_with("n_"), names_prefix="n_", names_to="race_spouse",
               values_to = "n") |>
  mutate(prop = n/total,
         race_ego = as.character(race_ego),
         spouse_multi = str_detect(race_spouse, "-"),
         spouse_const1 = str_split(race_spouse, "-", simplify = TRUE)[,1],
         spouse_const2 = str_split(race_spouse, "-", simplify = TRUE)[,2],
         overlap = case_when(
           race_ego == race_spouse ~ "Exact",
           !spouse_multi & str_detect(race_ego, race_spouse) ~ "Partial",
           spouse_multi & 
             (str_detect(race_ego, spouse_const1) | 
                str_detect(race_ego, spouse_const2)) ~ "Partial",
           TRUE ~ "None"
         ),
         overlap = factor(overlap,
                          levels=c("None", "Partial", "Exact")),
         race_spouse=factor(race_spouse, 
                            levels=c("White","Black","Indigenous","Asian",
                                     "Latino","White-Asian", "White-Indigenous",
                                     "White-Black", "Black-Asian",
                                     "Black-Indigenous", "Asian-Indigenous")),
         race_ego = factor(race_ego,
                           levels = c("White-Asian", "White-Indigenous", "White-Black",
                                      "Black-Asian", "Black-Indigenous", "Asian-Indigenous"))) |>
  ggplot(aes(x = fct_rev(race_spouse), y = prop, fill=overlap))+
  geom_col()+
  facet_wrap(~race_ego)+
  scale_fill_viridis_d(direction = -1, end=0.75)+
  scale_y_continuous(labels = scales::percent)+
  coord_flip()+
  labs(x="race of other spouse", y="percent", fill = "overlap in racial identification")+
  theme_bw()
```

@fig-partner-choice-panel-color also demonstrates that the probability of having a multiracial partner is relatively low across all six groups. However, the probability of having a multiracial partner rises substantially for each multiracial group when that multiracial partner is an exact match in terms of racial identification. This finding suggests some degree of exact endogamous preference for multiracial individuals.

Although @fig-partner-choice-panel-color provides a view of the overall experience of interracial marriage for multiracial individuals, its failure to account for group size, in addition to other factors, makes it difficult to understand the underlying social distance between groups. Thus, we turn to our model results to gain a better understanding of these patterns. All of the remaining results are based on the models described in the preceding section, which account for differences in group size in the local marriage market, as well as age, education, language, and birthplace differences between potential partners. Full results from these models are available in the supplementary materials.

Results are presented graphically by the relative odds, or odds ratio, of a given type of exogamous union relative to an exact endogamous union. For multiracial individuals, an endogamous union is signified by marriage between partners belonging to the same exact combination of races (e.g. a White-Black/White-Black union), which we refer to as “exact endogamy.” We also use color shading to distinguish cases of “partial endogamy” where multiracial partners have a partial overlap in racial identification with potential partners. We divide our discussion of the results between two types of unions, multiracial to multiracial unions and multiracial to monoracial unions.

For each point estimate of the odds ratio, we provide both 95% and 83.4% confidence intervals. Although commonplace in practice, 95% confidence intervals cannot reliably be used to determine if two point estimates are statistically distinguishable from each other at $p < 0.05$, although overlap in these intervals is sometimes incorrectly interpreted to indicate such. Lack of overlap in the 83.4% confidence intervals, on the other hand, is roughly equivalent to a confirmation that those point estimates are statistically distinguishable from each other at $p <0.05$.[^5] 

[^5]: This expectation only holds exactly when the standard errors for both point estimates are the same. See @knol_misuse_2011a for more details.

## Multiracial to Multiracial Intermarriage

@fig-odds-ratios-multi shows the odds relative to exact endogamy of multiracial individuals choosing a marriage partner from a different multiracial group (e.g. the odds a White-Black person marries a White-Indigenous partner relative to the odds of marrying another White-Black partner), across all fifteen possible combinations. To examine the role of partial overlap in racial identification, we distinguish estimates where spouses have a racial identification in common from those where no overlap is present. Due to small sample sizes, the standard errors of these estimates are substantial. We, nonetheless, draw two conclusions from the results. First, all of these odds ratios are moderate in size, casting doubt on the expectation that a shared experience of multiraciality leads to affinity across distinct multiracial combinations. The point estimates of the odds of multiracial-to-multiracial exogamy are at most 28.1% as high as those for exact endogamy, and the median and mean odds ratios are 11.1% and 12.3%, respectively. For comparison, these odds ratios are similar to the odds ratios of monoracial-to-monoracial exogamy which have a median and mean of 11.8% and 14.1%, respectively, and a maximum of 36.6% (White/Latino marriage). Thus, we find no evidence of more blurred boundaries between multiracial individuals belonging to different groups than among monoracial individuals belonging to different groups. The high likelihood of exact endogamy suggests that racial boundaries remain strong in the marriage choices of multiracial individuals.

```{r}
#| label: fig-odds-ratios-multi
#| fig-cap: Odds of interracial marriage with other multiracial groups for each multiracial group. 83.4% and 95% confidence intervals are shown for each estimate with thick and thin bars, respectively. The grey line indicates the case of an odds equal to one, meaning this type of intermarriage is as likely as endogamy. Bolding on labels indicates a shared racial identification.

coef_full |>
  mutate(race_ego=str_split(variable, "\\.", simplify=TRUE)[,1],
         race_choice=str_split(variable, "\\.", simplify=TRUE)[,2],
         race_cons1=str_split(race_ego, "-", simplify=TRUE)[,1],
         race_cons2=str_split(race_ego, "-", simplify=TRUE)[,2],
         overlap=str_detect(race_choice, race_cons1) | 
           str_detect(race_choice, race_cons2),
         overlap = factor(overlap, 
                              levels = c(FALSE, TRUE),
                              labels = c("No overlap", "Partial overlap")),
         variable=str_replace(variable, "\\.","/")) |>
  mutate(variable = if_else(str_count(variable, "White") == 2, 
                            str_replace_all(variable, "White", "**White**"), 
                            variable),
         variable = if_else(str_count(variable, "Black") == 2, 
                            str_replace_all(variable, "Black", "**Black**"), 
                            variable),
         variable = if_else(str_count(variable, "Indigenous") == 2, 
                            str_replace_all(variable, "Indigenous", "**Indigenous**"), 
                            variable),
         variable = if_else(str_count(variable, "Asian") == 2, 
                            str_replace_all(variable, "Asian", "**Asian**"), 
                            variable)) |>
  ggplot(aes(x=reorder(variable, coef, max), y=exp(coef), color=overlap))+
  geom_linerange(linewidth=1.25, 
                 aes(ymax=exp(coef+1.386*se), ymin=exp(coef-1.386*se)))+
  geom_linerange(linewidth=0.5, 
                 aes(ymax=exp(coef+1.96*se), ymin=exp(coef-1.96*se)))+
  geom_point(size=2)+
  geom_hline(yintercept=1, linewidth=0.5, color="grey30")+
  coord_flip()+
  scale_color_viridis_d(direction=-1, end=0.75)+
  labs(x=NULL, 
       y="odds of exogamous marriage with multiple race partner relative to multiracial endogamy",
       color="partial overlap")+
  theme_bw()+
  theme(axis.text.y = element_markdown())
```

Second, we find that the odds of multiracial-to-multiracial exogamy are generally higher when the two multiracial groups under consideration are partially endogamous. As @fig-odds-ratios-multi shows, the lowest odds of intermarriage were for the three cases in which partners had no shared racial identification in common (e.g. the Black-Indigenous group has no overlap with the White-Asian group). Thus, we find that “partial endogamy” as measured by a partial overlap in racial identification increases the odds of a union, although such unions are still far less likely than an exact match.

Among groups with a partial overlap in racial identification, @fig-odds-ratios-multi shows substantial variation in the odds of interracial marriage across groups. However, we can identify no particular pattern to account for this variation. Notably, the patterns observed here do not line up with the expectations of racial classification regimes. For example, we would have expected that shared Black identity would be stronger for multiracial partners and thus marriages which involved shared Black ancestry to be more likely than other multiracial-to-multiracial exogamous marriages. Similarly, we would have expected marriages involving shared Indigenous ancestry to be the least likely. Instead, the results here indicate no clear patterning to which shared ancestry is the most likely to lead to marriage between multiracial partners. However, due to small sample sizes, we cannot confidently say whether our failure to see such patterns is due to their non-existence or our lack of precision.

Given these findings, we estimate a more parsimonious model that collapses each specific multiracial-to-multiracial pairing into two dummy variables indicating unions with partial overlap in racial identification and unions with no such overlap. This model fits substantially better (a reduction of 54.3 on the BIC statistic) than the full model results and the two terms are statistically distinguishable from each other. According to this model, the odds of a union between mutiracial partners with partial overlap in racial identification are 17.2% as high as those for an exact endogamous union. The odds of a union between multiracial partners who share no racial identification are only 4.0% as high as those for an endogamous union. This model confirms that partial overlap in racial identification substantially increases the odds of interracial marriage between multiracial groups, but those odds are still substantially lower than the odds of exact endogamy. We apply this more parsimonious coding in all of the following models.

## Multiracial to Monoracial Intermarriage

@fig-odds-ratios-single shows the odds of someone from the focal multiracial group marrying a partner of a specific monoracial group, relative to the odds of exact endogamy (e.g. the odds of a White-Asian person marrying a White partner rather than a White-Asian partner). Consistent with the results for multiracial-to-multiracial intermarriage, we find that the tendency toward exact endogamy is quite strong. Almost all of the odds ratios here are below 0.25, indicating that the odds of a union with a monoracial person, even when sharing partial racial identification, is less than 25% as high as the odds of marrying a multiracial person of the same exact racial combination. The strength of this exact endogamy was masked in @fig-partner-choice-panel-color by the small size of multiracial groups.

```{r}
#| label: fig-odds-ratios-single
#| fig-cap: Odds of interracial marriage with single race groups for each multiracial group. 83.4% and 95% confidence intervals are shown for each estimate with thick and thin bars, respectively. The grey line indicates an odds equal to one, meaning this type of interracial marriage is as likely as endogamy.

# the somewhat odd regular expression \\..*- will identify all the names with a 
# period followed by some characters followed by a dash. This will give
# me all cases of a single race/multiple race marriage
coef_short_sym |>
  filter(str_detect(variable, "\\..*-")) |>
  mutate(single_race = str_split(variable, "\\.", simplify=TRUE)[,1],
         multiple_race = str_split(variable, "\\.", simplify=TRUE)[,2],
         constituent = str_detect(multiple_race, single_race),
         single_race = factor(single_race, 
                            levels=c("Latino","Asian","Indigenous",
                                     "Black","White")),
         multiple_race = factor(multiple_race,
                              levels=c("White-Asian","Black-Asian",
                                       "White-Indigenous","Black-Indigenous",
                                       "White-Black","Asian-Indigenous")),
         constituent = factor(constituent, 
                              levels = c(FALSE, TRUE),
                              labels = c("None", "Partial")),
         upper84=exp(coef+1.386*se),
         upper95=exp(coef+1.96*se),
         lower84=exp(coef-1.386*se),
         lower95=exp(coef-1.96*se)) |>
  ggplot(aes(x=single_race, y=exp(coef), color=constituent))+
  geom_linerange(linewidth=1.25, 
                  aes(ymax=upper84, ymin=lower84))+
  geom_linerange(linewidth=0.5, 
                  aes(ymax=upper95, ymin=lower95))+
  geom_point(size=2)+
  geom_hline(yintercept=1, linewidth=1, color="grey30")+
  coord_flip()+
  facet_wrap(~multiple_race, ncol=2)+
  scale_color_viridis_d(direction=-1, end=0.75)+
  labs(x="single race partner", 
       y="odds of marriage with given single race partner relative to endogamy",
       color = "overlap in racial identification")+
  theme_bw()
```

@fig-odds-ratios-multi also distinguishes between monoracial groups that are constituent of a given multiracial category and those that are not. Putting aside for the moment the results for Latino partners, the odds of interracial marriage are generally higher when the monoracial group is constituent of the multiracial category, indicating that partial overlap in racial identification increases the likelihood of union formation.

These estimates are also consistent with our predictions based on racial classification regimes. Consistent with the *Black classification regime*, the odds of marrying a Black partner are higher than the odds of marrying a partner of the other constituent racial group for all three cases involving a part-Black multiracial group. For example, the odds of a White-Black multiracial person marrying a Black spouse are higher than the odds of marrying a White spouse (OR=0.253 vs. OR=0.179). Conversely, as the *Indigenous classification regime* would predict, the odds of marrying an Indigenous partner are substantially lower than the odds of marrying the other component group for all three cases involving a part-Indigenous multiracial group. For example, the odds of a White-Indigenous person marrying a White spouse are substantially higher than the odds of marrying an Indigenous spouse (OR=0.236 vs. OR=0.066). These patterns suggest that dynamics of racial dominance and recession for classification of multiracial groups also have bearing on how multiracial people are treated on the marriage market, at least with respect to potential monoracial partners.

Part-Asian groups have intermarriage patterns that reflect an “in-between” racial regime. The results for Black-Asian and Asian-Indigenous individuals are somewhat imprecise due to smaller sample sizes, but are generally consistent with the expectations of the Black and Indigenous classification regimes, respectively. White-Asian individuals, for whom we have less clear expectations, are the most likely of any multiracial group to partner with both monoracial component groups. They are slightly more likely to marry White partners than Asian partners, but the difference is not statistically significant.

## Latinos and Multiracial Intermarriage

We observe that all six multiracial groups have a substantial likelihood of intermarriage with Latinos. In many cases, these odds are close to or even higher than the odds of marrying a monoracial group with a partial overlap in racial identification. This finding may be explained by broader patterns of Latino intermarriage. We illustrate these patterns in @fig-odds-ratios-latino, which shows the overall odds of interracial marriage for Latinos to all other groups, relative to endogamy.

```{r}
#| label: fig-odds-ratios-latino
#| fig-cap: Odds of interracial marriage for Latinos with single and multiracial race partners. The grey line indicates the case of an odds equal to one, meaning this type of intermarriage is as likely as endogamy.
#| 
coef_short_sym |>
  filter(str_detect(variable, "Latino")) |>
  mutate(
    variable=str_replace(variable, "Latino",""),
    variable=str_replace(variable, "\\.", ""),
    type=case_when(
      str_count(variable, "-")>0 ~ "Multiracial partner",
      TRUE ~ "Monoracial partner"),
    upper84=exp(coef+1.386*se),
    upper95=exp(coef+1.96*se),
    lower84=exp(coef-1.386*se),
    lower95=exp(coef-1.96*se)) |>
  filter(variable!="") |>
  ggplot(aes(x=reorder(variable, coef, max), y=exp(coef), color=type))+
  geom_hline(yintercept=1, linewidth=1, color="grey30")+
  geom_point()+
  geom_linerange(linewidth=1.25, 
                  aes(ymax=upper84, ymin=lower84))+
  geom_linerange(linewidth=0.5, 
                  aes(ymax=upper95, ymin=lower95))+
  scale_color_viridis_d(direction=-1, end=0.75)+
  labs(x="Race of non-Latino partner",
       y="odds of intermarriage for Latino partners relative to endogamy",
    color="non-Latino partner")+
  coord_flip()+
  theme_bw()
```

The highest odds of interracial marriage for Latinos is with White partners. The odds here are 36.6% as high as those for Latino endogamy. This odds ratio is the largest of any of the racial exogamy terms we estimate in the models. The next three groups with the highest likelihood of interracial marriage with Latinos are the three part-White multiracial groups. At the other end of the spectrum, Latinos are the least likely to marry Black individuals and part-Black multiracial groups without White identification.

Although Latinos are least likely to marry monoracial Black partners of all the single-race options, they still have a higher odds of marrying Black partners than single race White, Asian, and Indigenous individuals (see full model table in supplementary materials). Thus, we believe two factors account for the high odds of intermarriage between Latinos and multiracial spouses. First, Latinos have higher rates of interracial marriage than other groups in general. Second, this potential for interracial marriage is particularly high for potential White partners, which carries over to multiracial groups with part-White racial identification. Thus, Latinos might be characterized as more “open” to interracial marriage even as they tend to conform to a racial hierarchy with Whiteness at the top and Blackness at the bottom. One notable caveat to this pattern is the relatively high odds of intermarriage with White-Black individuals which would seem to contradict the expectations of the Black racial classification regime.


## Multiracial Intermarriage and Gender: The Case of White-Black and White-Asian Partner Choices

As a final test of the power of racial regimes in the partnering decisions of multiracial individuals, we estimate a model in which we allow for gender asymmetry in the effects of racial exogamy. We use these results to look at the gender asymmetry into marriage for White-Black and White-Asian individuals.

Researchers have long noted a gender asymmetry in the likelihood of interracial marriage for Black and Asian partners. Black men are more likely to marry across racial lines than Black women, while Asian women are more likely to marry across racial lines than Asian men [@jacobs_gender_2002a]. These differences may reflect the way race and gender interact in the evaluation of potential partners [@curington_romantic_2022]. 

If part-Black individuals are more likely to be treated as Black on the marriage market, we would expect to observe the same gender asymmetry as for Black individuals. Similarly, if part-Asian individuals are more likely to be treated as Asian, we would expect the same gender asymmetry as for Asian individuals. To test this formally, @fig-gender-asymmetric-focused shows the gender-specific odds of intermarriage with single race White partners for White-Black and White-Asian spouses, in comparison to the same results for Black and Asian spouses, respectively. We focus on these groups because the sample sizes are large enough to more adequately test for such differences.[^gender-figs]

[^gender-figs]: We provide a figure showing the gender specific results across all monoracial groups in the supplementary materials. This figure is generally consistent with our findings for gender asymmetry with White partners, but some of the results are less certain due to larger standard errors.

```{r}
#| label: fig-gender-asymmetric-focused
#| fig-cap: Odds of interracial marriage with a White partner separately by gender for White/Asian, Asian, White/Black, and Black spouses. 83.4% and 95% confidence intervals are shown for each estimate with thick and thin bars, respectively.

# before period is husband, after period is wife

temp <- coef_short_asym |>
  filter(str_detect(variable, "^Black\\.|\\.Black$") &
           !str_detect(variable, "-") &
           !str_detect(variable, "Multi/Multi|Latino")) |>
  mutate(race_husband=str_split(variable, "\\.", simplify=TRUE)[,1],
         race_wife=str_split(variable, "\\.", simplify=TRUE)[,2],
         ego_race = "Black",
         alt_race = ifelse(str_detect(race_husband, "Black"), race_wife,
                            race_husband),
         race_label=ifelse(str_detect(race_husband, "Black"), 
                           paste(race_wife, " Female", 
                                 sep=""),
                           paste(race_husband, " Male", 
                                 sep="")))

temp <- coef_short_asym |>
  filter(str_detect(variable, "^Asian\\.|\\.Asian$") &
           !str_detect(variable, "-") &
           !str_detect(variable, "Multi/Multi|Latino")) |>
  mutate(race_husband=str_split(variable, "\\.", simplify=TRUE)[,1],
         race_wife=str_split(variable, "\\.", simplify=TRUE)[,2],
         ego_race = "Asian",
         alt_race = ifelse(str_detect(race_husband, "Asian"), race_wife,
                            race_husband),
         race_label=ifelse(str_detect(race_husband, "Asian"), 
                           paste(race_wife, " Female", 
                                 sep=""),
                           paste(race_husband, " Male", 
                                 sep=""))) |>
  bind_rows(temp)

temp <- coef_short_asym |>
  filter(str_detect(variable, "^White-Black\\.|\\.White-Black$") &
           !str_detect(variable, "Multi/Multi|Latino")) |>
  mutate(race_husband=str_split(variable, "\\.", simplify=TRUE)[,1],
         race_wife=str_split(variable, "\\.", simplify=TRUE)[,2],
         ego_race = "White-Black",
         alt_race = ifelse(str_detect(race_husband, "White-Black"), race_wife,
                            race_husband),
         race_label=ifelse(str_detect(race_husband, "White-Black"), 
                           paste(race_wife, " Female", 
                                 sep=""),
                           paste(race_husband, " Male", 
                                 sep=""))) |>
  bind_rows(temp)

temp <- coef_short_asym |>
  filter(str_detect(variable, "^White-Asian\\.|\\.White-Asian$") &
           !str_detect(variable, "Multi/Multi|Latino")) |>
  mutate(race_husband=str_split(variable, "\\.", simplify=TRUE)[,1],
         race_wife=str_split(variable, "\\.", simplify=TRUE)[,2],
         ego_race = "White-Asian",
         alt_race = ifelse(str_detect(race_husband, "White-Asian"), race_wife,
                            race_husband),
         race_label=ifelse(str_detect(race_husband, "White-Asian"), 
                           paste(race_wife, " Female", 
                                 sep=""),
                           paste(race_husband, " Male", 
                                 sep=""))) |>
  bind_rows(temp)

temp <- temp |>
  filter(alt_race == "White") |>
  mutate(
    group = case_when(
      ego_race %in% c("White-Asian", "Asian") ~ "Asian or White-Asian ego partner",
      ego_race %in% c("White-Black", "Black") ~ "Black or White-Black ego partner"
    ), 
    group = factor(group,
                   levels = c("Black or White-Black ego partner",
                              "Asian or White-Asian ego partner")),
    race_type = factor(str_detect(ego_race, "-"),
                       levels = c(FALSE, TRUE),
                       labels = c("Monoracial", "Multiracial"))
  ) |>
  select(variable, coef, se, group, race_type, race_label)

temp <- temp |>
  mutate(upper84=exp(coef+1.386*se),
         upper95=exp(coef+1.96*se),
         lower84=exp(coef-1.386*se),
         lower95=exp(coef-1.96*se))

temp |>
  ggplot(aes(x=factor(race_label), y=exp(coef), color=race_type))+
  #geom_hline(yintercept=1, linewidth=0.5, color="grey30")+
  geom_linerange(linewidth=1.25, 
                 aes(ymax=upper84, ymin=lower84))+
  geom_linerange(linewidth=0.5, 
                 aes(ymax=upper95, ymin=lower95))+
  geom_point(size=2)+
  coord_flip()+
  facet_wrap(~group, ncol=1)+
  scale_color_viridis_d(direction=-1, end=0.9)+
  theme_bw()+
  labs(color="ego partner race",
       x="other partner race and gender", 
       y="odds of marriage with White partner relative to endogamy")
```

In the case of White-Black partners, we observe the same basic gender asymmetry observed for Black partners. The results for monoracial Black partners show that the odds of a Black man/White woman pairing are substantially more likely than a Black woman/White man pairing. The overall odds of intermarriage with White partners are substantially higher for White-Black individuals than for Black individuals, but maintain the same basic gender asymmetry. White-Black man/White woman pairings are more likely than White-Black woman/White man pairings. These findings reinforce the earlier findings that the Black classification regime structures the marital choices and opportunities of part-Black individuals, whose marriage patterns more closely resemble those of single-race Black individuals than their other component race.

In contrast, we find no evidence of this kind of gender asymmetry for White-Asian partners. @fig-gender-asymmetric-focused shows that the odds of marriage are much stronger for Asian woman/White man pairings than for Asian man/White woman pairings, as expected. However, the odds of a White-Asian/White marriage are similar for White-Asian men and women. While the point estimates suggest a slight gap in the expected direction, this difference is not statistically distinguishable from zero. Overall, consistent with the earlier results, the results here suggest more ambiguity in the racial classification regime for Asian ancestry than for Black ancestry.

# Conclusions

Marriages involving multiracial individuals can deepen our understanding of how interracial partnering shapes racial boundaries by focusing on partnering occurring in the “next generation down” [@song_what_2015a, p. 95]. Our results indicate a level of exact endogamy among multiracial individuals sharing the same racial combination that is similar in magnitude to the levels of endogamy observed for monoracial groups. Additionally, we find that “partial” endogamy where potential monoracial and multiracial partners have a partial overlap in racial identification also substantially increases the likelihood of union formation. In sharp contrast, we find a very low likelihood of union formation when multiracial partners do not share at least a partial racial identification with potential partners. Thus, we find no support for the hypothesis that multiracial partners sort into marriages based on a general shared experience of multiraciality. Furthermore, we find no evidence that multiracial individuals marry in a postracial or transcendent fashion, in which race plays little or no role in their partner choices. Instead, we find that while the racial boundaries around multiracial people may be more complex than those of monoracial people, these boundaries still strongly determine partner selection.

Our results may overstate the tendency toward exact endogamy among multiracial individuals because we measure racial identification after rather than before marriage formation and partners may switch to more similar identifications after the marriage. However, given the relative strength of endogamy in these results, we do not believe such a bias could be sufficiently large enough to invalidate our overall findings.

We also find that partial overlap in racial identification has variable and patterned effects on the likelihood of union formation, depending on which race overlaps. Specifically, we find that these patterns correspond to historic racial classification regimes that also determine the dominant racial identification of multiracial individuals [@iverson_regimes_2022a]. These patterns reflect the tendencies in identity formation to inhabit single-race categories, through strategic performances of one’s identity [@khanna_if_2010a] as well as a felt closeness to specific racial groups [@jones_making_2023a; @davenport_politics_2018a]. Shared Black ancestry is most predictive of union formation, while shared Indigenous ancestry is least predictive, and shared White ancestry falls somewhere in-between. Shared Asian ancestry operates somewhat like White ancestry in responding to the dominance of Black ancestry and the recessiveness of Indigenous ancestry and is co-equal in dominance with White ancestry. These findings suggest a less established classification regime for Asian ancestry, consistent with other work [@gullickson_choosing_2011a; @iverson_regimes_2022a]. In general, the patterns predicted by racial classification regimes are more apparent when we examine multiracial-to-monoracial intermarriage than multiracial-to-multiracial intermarriage. While this difference may reflect a weaker role of racial classification regimes when both partners are themselves multiracial, it may also be a result of more statistical noise in a smaller sample.

Substantial levels of endogamy are believed to contribute to stronger and more salient racial boundaries and a greater sense of groupness [@schwartz_trends_2013a]. The tendency towards endogamy among multiracial groups may reflect a growing sense of groupness for combination-specific multiracial groups [@dacosta_making_2007a; @dacosta_multiracial_2020a; @jones_who_2011a]. However, we consider this scenario unlikely because the tendency toward endogamy among multiracial groups is largely counteracted by the small size of these groups. Even if multiracial individuals might prefer to marry individuals of the exact same multiracial category, they are most likely to marry individuals of a single race because they have many more available options from among these larger populations. Moreover, because the largest single race group is the White group and the largest multiracial groups have part-White ancestry, many multiracial individuals will overlap specifically with a single-race White partner.

This group size issue has important implications for how we think about the future of multiraciality in the US. Multiracial groups are most likely to partner with monoracial groups with whom they share a partial racial identification. @bratter_will_2007a has shown that the children of such unions are often identified by the single shared race of the partners. Therefore, when viewed as a multigenerational process, the potential of a growing self-identified multiracial population to redefine racial boundaries will be tempered by the tendency for family genealogies to “fold” back into prominent single race categories across multiple generations. Due to the demographic realities discussed above, the most common single race category to be folded back into will be the White category.

A further important implication of this research is the relatively high rate of intermarriage between Latinos and non-Latino multiracial groups. This finding reflects to some extent a higher rate of exogamy among Latinos more generally. Nonetheless, this pattern is noteworthy for combining demographic groups who, for different reasons, are often racially ambiguous within the US context [@fox_defining_2013a]. The complex ancestry arising from such combinations may further complicate Latino identity and classification in the US context [@miyawaki_who_2024a].

There are important limitations to this research. Our approach excludes those who have racially mixed ancestry but opt not to identify with more than one race. Ultimately, all analyses of multiracial populations are bounded by a focus on “... *a* multiracial population, not *the* multiracial population” [@harris_who_2002a, p. 625]. Our decision to focus on self-identification allows us to capture the most self-aware population who are most likely to transmit their ancestral identification onto the next generation [@bratter_will_2007a; @song_multiracial_2016a], which complements our interest in the intergenerational process of multiracial identification. However, by definition, this approach excludes others whose sense of multiracial identification may be weaker.

This analysis is also limited to marital unions only, due to limitations of data and space. However, all forms of interracial partnering are important to understand, particularly if we are interested in multigenerational processes. Future work on potential differences in the patterns found here across partnering types would be valuable in this regard.

Finally, we are not able to include multiracial individuals of part-Latino ancestry because of question design on the ACS. While we do note that Latinos have particularly high rates of intermarriage with other multiracial groups, we are not able to capture the specific experiences of multiracial/multiethnic Latinos in the marriage market. Recent changes approved by the federal government [@officeofmanagementandbudget_revisions_2024a] should allow future data to make this distinction, presenting an opportunity for future research to better identify part-Latino individuals, and provide a fuller picture of the overall experience of multiraciality in the United States.
